---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# mapler

<!-- badges: start -->
[![R-CMD-check](https://github.com/matthewwhite1/mapler/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/matthewwhite1/mapler/actions/workflows/R-CMD-check.yaml)
<!-- badges: end -->

The goal of mapler is to provide functions to investigate how maple
syrup production could be impacted by climate change.

## Installation

You can install the development version of mapler from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("matthewwhite1/mapler")
```

## Downloading Climate Data

This package provides three functions for downloading climate data: `download_ghcnd()`,
`download_loca2()`, and `download_prism()`. These functions are useful for downloading
both historic climate data and future climate projections.

Once data is downloaded, it can be loaded into R using `loca_t_rast()` or
`prism_t_rast()`, "t" meaning temperature. These functions take filepaths
of where the data is downloaded and return the tmin and tmax raster stacks.

## Calculating Sap Days

Maple sap tapping is highly dependent on the freeze/thaw cycle. This package calls
a "sap day" a day where the temperature switched between freezing and thawing
during the day.

The function `sap_day()` takes in tmin and tmax raster stacks and aggregates them
into a yearly sap day stack. This combination of tmin and tmax is used to calculate
the proportion of sap days in a year at every pixel. This function
can be used for future climate projection with LOCA2 data.

The `sap_day()` function can be used with LOCA2 data like so:

```{r sap_day, eval = FALSE}
# Load in rasters
loca_rast <- loca_t_rast("F:/Data/LOCA2/ACCESS-CM2/0p0625deg/r1i1p1f1/")

# Calculate yearly sap days (LOCA2 temps are in Kelvin)
k_upper <- 2.2 + 273.15
k_lower <- -1.1 + 273.15
loca_sap_day <- sap_day(loca_rast$tmax, loca_rast$tmin,
                        t_upper = k_upper, t_lower = k_lower)

# Shift raster to correct longitude to match WGS84
test_loca <- loca_sap_day$proportion |>
  terra::shift(dx = -360)
```

## Package Data

This package contains a few datasets from various sources. The `farms` datasets
(`us_farms`, `canada_farms`, `quebec_farms`) contain maple farm names and
addresses scraped from the internet. The `farms_coords` dataset contains
geocoded latitude and longitude coordinates of these farms.

This package also contains two datasets to assist in downloading climate data:
`loca2_model_names` and `ghcnd_stations`.

For an example of what `sap_day()` can do, this package contains an example 
raster called `test_loca_sap_day`. This was created by downloading historical
and scenario ssp585 tmin and tmax data from the LOCA2 model ACCESS-CM2,
loading the rasters into R using `loca_t_rast()`, calculating yearly sap
day proportions using `sap_day()`, shifting the longitude by -360 to
match WGS84, and aggregating by a factor of 20 to decrease the file size. It
can be accessed like so:

```{r test_loca}
test_loca_file <- system.file("extdata", "test_loca_sap_day.tif",
                                package = "mapler")
sap_prop <- terra::rast(test_loca_file)
```

## Sen's Slope

This package contains a function called `sens_slope_rast()` that calculates
the Sen's slope for every pixel in a raster stack. This function is based on
the Rcpp-optimized `sens_slope()` function, which can be used in contexts
separate from maple syrup. These functions can be used like so:

```{r sens_slope}
library(mapler)
library(terra)

# Get Sen's slope of vector of random values
x <- rnorm(1000)
sens_slope(x)

# Get and plot Sen's slope of entire raster
test_loca_file <- system.file("extdata", "test_loca_sap_day.tif",
                                package = "mapler")
sap_prop <- terra::rast(test_loca_file)
sens_rast <- sens_slope_rast(sap_prop)
plot(sens_rast[[1]])
```

## Future Climate Projection

The `sens` family of functions can be used for future climate projection.
Output from `sap_day()` can be passed into `sens_farms()`, along with farms
coordinates. This function combines the `get_sens_farms()` and
`get_sens_joined()` functions. `sens_farms()` starts by calculating the 
Sen's slope for each farm and creating a new variable that is 1 if the slope is
significant positive, -1 if the slope is significant negative, and 0 otherwise.
Then, the farms are joined to a shapefile, and a proportion of the significant
variable is calculated for each group in the shapefile. This resulting dataframe
can be plotted with something like ggplot.

The following code can be used as a guide for plotting a future sap day projection.

```{r eco_region, message = FALSE, warning = FALSE, fig.dim = c(10, 8)}
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)
library(tidyverse)
library(RColorBrewer)

# Read in farm coordinates and sap day projection
farms_sf <- sf::st_as_sf(farms_coords, coords = c("lon", "lat"), crs = 4326)
test_loca_file <- system.file("extdata", "test_loca_sap_day.tif",
                                package = "mapler")
sap_prop <- terra::rast(test_loca_file)

# Read in eco regions shape file
shapefile <- sf::read_sf("Data_Clean/NA_Eco_Level3/NA_CEC_Eco_Level3.shp")
variable <- names(shapefile)[2]

# Get proportion of Sen's significance at each eco region
eco_regions_joined <- sens_farms(farms_coords = farms_sf, sap_prop = sap_prop,
                                 shapefile = shapefile, group_var = variable)

# Prepare map for plotting
world <- ne_countries(scale = "medium", returnclass = "sf")
north_america <- world |>
  filter(region_un == "Americas", name %in% c("United States of America", "Canada"))
us_states <- ne_states(country = "United States of America", returnclass = "sf")
canada_provinces <- ne_states(country = "Canada", returnclass = "sf")

# Prepare colors
breaks <- seq(-1, 1, by = 0.2)
eco_regions_joined$sig_mean_class <- cut(eco_regions_joined$sig_mean, 
                                         breaks, include.lowest = TRUE)

# Plot eco regions colored by significance proportion
ggplot() +
  geom_sf(data = north_america, fill = "grey85", color = "black", size = 0.2, alpha = 0.7) +
  geom_sf(data = us_states, fill = NA, color = "grey40", size = 0.3) +
  geom_sf(data = canada_provinces, fill = NA, color = "grey40", size = 0.3) +
  geom_sf(data = eco_regions_joined, mapping = aes(fill = sig_mean_class)) +
  geom_sf(data = farms_sf, color = "black", size = 1.5) +
  coord_sf(xlim = c(-99, -59), ylim = c(32, 52), expand = FALSE) +
  scale_fill_manual("Significance Proportion", 
                    values = rev(brewer.pal(10, "PuOr"))[c(1, 3, 5, 6, 7, 10)]) +
   theme_bw() +
  labs(
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(legend.position = "inside",
        legend.position.inside = c(0.85, 0.27),
        text = element_text(size = 16),
        legend.text = element_text(size = 10),
        legend.title = element_text(size = 10),
        panel.grid.major = element_line(linewidth = 1))
```

The same thing can be done for U.S. counties:

```{r us_counties, message = FALSE, warning = FALSE, fig.dim = c(10, 8)}
library(tigris)

shapefile <- counties(cb = TRUE, year = 2020)
variable <- "GEOID"

us_counties_joined <- sens_farms(farms_coords = farms_sf, sap_prop = sap_prop,
                                 shapefile = shapefile, group_var = variable)

# Prepare colors
breaks <- seq(-1, 1, by = 0.2)
us_counties_joined$sig_mean_class <- cut(us_counties_joined$sig_mean, 
                                         breaks, include.lowest = TRUE)

# Plot
ggplot() +
  geom_sf(data = north_america, fill = "grey95", color = "black", size = 0.2) +
  geom_sf(data = us_states, fill = NA, color = "darkgray", size = 0.3) +
  geom_sf(data = canada_provinces, fill = NA, color = "darkgray", size = 0.3) +
  geom_sf(data = us_counties_joined, mapping = aes(fill = sig_mean_class)) +
  geom_sf(data = farms_sf, color = "black", size = 0.5) +
  coord_sf(xlim = c(-99, -59), ylim = c(32, 53), expand = FALSE) +
  scale_fill_manual("Significance Proportion", 
                    values = rev(RColorBrewer::brewer.pal(5, "OrRd"))) +
  theme_minimal() +
  labs(
    # title = "Significance of Sens Slope at Maple Farms",
    x = "Longitude",
    y = "Latitude"
  ) +
  theme(legend.position = "inside",
        legend.position.inside = c(0.9, 0.22))
```
